Document knowledge base linking for information retrieval

### 2016-02-11 paper.tex
revisions

In this paper, we present a method of using standard information retrieval techniques to produce knowledge base links for a corpus of documents. Our approach benefits by removing the requirement that links connect specific mentions from the text. We argue that such a ``bag of links'' provides many of the same benefits as more fine-grained entity linking or wikification processes for the purposes of ad hoc document retrieval. To demonstrate this benefit, we employ our knowledge base links using a simple document expansion model, which yields improvement over IR baselines.

@@

Entity linking and wikification have been studied extensively, particularly in the context of TAC KBP \cite{Ji2014}. Variations on these tasks exist: in some instances, entity mentions are assumed to be available, essentially reducing the tasks to disambiguation; more recently, the tasks includes identification of mentions in raw document text. If a system must identify entity mentions, it is common to use a named entity recognition tagger \cite{Hoffart2011, Ratinov2011}

Entity linking systems often exploit knowledge base structure to help match and disambiguate entities. For example, Cucerzan \cite{Cucerzan2007} employs Wikipedia disambiguation pages and redirection pages to help identify various entity surface forms. The use of anchor text and hyperlink graphs is also a frequent feature of these systems \cite{Mihalcea2007, Ratinov2011, Stoyanov2012}. Most systems employ some form of context, which often refers to the co-occurrence of entities as evidence for disambiguation \cite{Stoyanov2012, Dalton2013}. Another use of context, relating most closely to our approach in this paper, refers the text surrounding an entity mention, which can be used to disambiguate knowledge base representations \cite{Mihalcea2007, Dalton2013}.

@@

\subsection{Document Expansion in Information Retrieval}\label{section.related.ir}

Investigation into the use of knowledge base links for document retrieval have, to our knowledge, been limited. Those applications most similar to ours use entity links in a feedback process to expand queries. Dalton and Dietz \cite{Dalton2014} use a number of sources, including entity-annotated queries and pseudo-relevance feedback (PRF) on entity-annotated document collections. Their process is underpinned by entity links produced by their KB Bridge system \cite{Dalton2013}. Similarly, Xiong and Callan expand queries using PRF with the descriptions and categories of Freebase objects linked to a retrieved document in the FACC1\footnote{http://lemurproject.org/clueweb09/FACC1/} annotation data. They employ a supervised machine learning model to select expansion terms, trained on the basis of the terms' influence on the ranking scores of relevant documents \cite{Xiong2015}.

@@

\begin{flalign}\label{equation.ql}
\hat{P}(w | \theta_D) = \frac{c(w, D) + \mu \hat{P}(w | C)}{|D| + \mu} 
\end{flalign}

@@

\subsection{Linking with Document Pseudo-Queries}\label{section.linking.queries}
%The KL-divergence retrieval model measures the distance between the query language model, $\theta_{D_Q}$ here, and the document language model, $\theta_E$, as follows:
%
%\begin{flalign*}
%	score(E,D_Q) 	&= -KL(\theta_{D_Q}||\theta_E) \\
%					&\rankeq \sum_{w \in V} P(w|\theta_{D_Q}) \log P(w|\theta_E)
%&\end{flalign*}
%
\noindent To find candidate entities to ``link" to a given document $D$, we begin by treating the text of $D$ as a pseudo-query which we pose against a collection of entities $C_E$.  To transform a document into a pseudo-query we apply two transformations.  First we remove all terms from $D$ that appear in the standard Indri stoplist\footnote{http://www.lemurpro ject.org/stopwords/stoplist.dft}.  Next, we prune our pseudo-query by retaining only the $0 < k \leq |D|$ most frequent words in the stemmed text of $D$.  The integer variable $k$ is a parameter that we choose empirically.  Let $Q_D$ be the pseudo-query for $D$, consisting of the text of $D$ after our two transformations.

@@

We now have a ranked list of tuples $\{(E_1, \pi_1)$, $(E_2, \pi_2)$, $...$, $(E_N, \pi_N)\}$ relating knowledge base entry $E_i$ to $D$ with log-probability $\pi_i$. Our final step is to take the top $n$ entries where $0 \leq n \leq N$. We call these top entries $\mathcal{E}_D$ and designate them as our knowledge base links for $D$.  Finally, we exponentiate each $\pi_i$ and normalize our entity scores so they sum to 1 over the $n$ retained entities.  Assuming a uniform prior over entities, we now have a probability distribution over our $n$ retained entities: $P(E | D)$.

@@

We test our approach using the TREC 2004 robust topics. These 250 topics are used with data from TREC disks 4 and 5. In addition, we use the AP newswire collection from TREC disks 1 and 2 with topics 101-200. % needs work

@@

We produce two runs per collection:
\begin{itemize}
	\item \textit{baseline-ql}, a baseline query likelihood run
%	\item \textit{baseline-rm3}, a baseline RM3 run
	\item \textit{kb-ql}, incorporating knowledge base links using Eq \ref{eq.ql-and-entities}.
%	\item \textit{kb-rm3}, incorporating knowledge base links using an initial RM3 retrieval
\end{itemize}

For the two \textit{kb} runs, we retrieve the top 1000 documents per query based on the default Indri query likelihood implementation. We then re-rank these documents by incorporating their knowledge base links as described in Section \ref{section.model}.

@@

\begin{table}[htbp]
\centering
\begin{tabular}{|c|l|c|c|} \hline
& {\bf Run} & {\bf $\lambda$} & {\bf MAP} \\ \hline
\multirow{2}{*}{Robust} & {\it baseline-ql} & -- & 0.2183 \\ \cline{2-4}
%& {\it baseline-rm3} & -- & 0.2548 \\ \cline{2-4}
& {\it kb-ql} & 0.9 & {\bf 0.2275} \\ \hline \hline %\cline{2-4}
%& {\it kb-rm3} & & \\ \hline \hline
\multirow{2}{*}{AP} & {\it baseline-ql} & -- & 0.2346 \\ \cline{2-4}
%& {\it baseline-rm3} & -- & 0.2653 \\ \cline{2-4}
& {\it kb-ql} & 0.9 & {\bf 0.2515} \\ \hline %\cline{2-4}
%& {\it kb-rm3} & 0.8 & 0.2665 \\ \hline
\end{tabular}
\caption{The top-scoring knowledge base runs and baselines by mean average precision (MAP). $\lambda$ is specified for knowledge base runs. Bolded values indicate statistical significance compared to the baseline.}
\label{table.performance}
\end{table}

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{figures/sweep-ql.pdf}
\caption{Sweeps over values of $\lambda$ for robust (red) and AP (blue) runs.}
\label{figure.sweeps-ql}
\end{figure}

\begin{comment}
\begin{figure}
\centering
\includegraphics[width=\columnwidth]{figures/sweep-rm3.pdf}
\caption{Sweeps over values of $\lambda$ for robust (red) and AP (blue) RM3 runs.}
\label{figure.sweeps-rm3}
\end{figure}
\end{comment}

Retrieval performance of the top-scoring knowledge base runs and baselines are shown in Table \ref{table.performance}. Bolded mean average precision (MAP) scores are greater than their corresponding baseline (e.g. \textit{kb-ql} compared to \textit{baseline-ql}) with statistical significance at $p < 0.001$ using a paired t-test. Note that baselines correspond to $\lambda = 0.0$.

Though Table \ref{table.performance} shows only the top-scoring values of $\lambda$, performance for all knowledge base runs improved over the baseline for values of $0.0 < \lambda < 1.0$, as shown in Figure \ref{figure.sweeps-ql}. All improvements were statistically significant with at least $p < 0.05$.

% Paragraph about RM3 results...

@@

The results indicate that our approach for constructing knowledge base links between documents and Wikipedia produces useful data for document retrieval purposes. Our simple document expansion model that incorporates these links performs well compared to a query likelihood baseline. These outcomes support our argument that a ``bag of links'' to a knowledge base can provide helpful information for a document retrieval task. Our proposed method using standard information retrieval techniques to build these links is efficient compared to usual entity linking and wikification systems and can be accomplished without the need for complex natural language processing.

>>>>

For the purposes of ad hoc document retrieval, however, the granularity of linking entities to specific mention spans in the text is not necessarily required. Since retrieval is concerned with scoring documents in their entirety, we may simplify the task from mapping \textit{mentions} to entities, as is more common, to mapping \textit{documents} to entities. We argue that such a ``bag of links'' provides many of the same benefits as more fine-grained entity linking or wikification processes for the purposes of ad hoc document retrieval while improving computational efficiency.

In this paper, we present a method of using standard information retrieval techniques to produce a ``bag of links'' for each document in a corpus. To demonstrate their utility, we employ our knowledge base links using a simple document expansion model, which yields improvement over IR baselines.

@@

Entity linking and wikification have been studied extensively, particularly in the context of TAC KBP \cite{Ji2014}. Entity linking systems often exploit knowledge base structure to help match and disambiguate entities. For example, Cucerzan \cite{Cucerzan2007} employs Wikipedia disambiguation pages and redirection pages to help identify various entity surface forms. The use of anchor text and hyperlink graphs is also a frequent feature of these systems \cite{Mihalcea2007, Ratinov2011, Stoyanov2012}. Most systems employ some form of context, which often refers to the co-occurrence of entities as evidence for disambiguation \cite{Stoyanov2012, Dalton2013}. Another use of context, relating most closely to our approach in this paper, refers the text surrounding an entity mention, which can be used to disambiguate knowledge base entries \cite{Mihalcea2007, Dalton2013}.

@@

\subsection{Document Expansion in IR}\label{section.related.ir}

Prior applications of knowledge base links in information retrieval use entity links in a feedback process for query expansion. Xiong and Croft, for example, employ FACC1\footnote{http://lemurproject.org/clueweb09/FACC1/} annotations on the document collection with supervised learning to expand queries \cite{Xiong2015}. Other research has used entity links in the queries themselves, rather than documents, to aid in expansion \cite{Xu2009, Brandao2014}. Still others use entity links in \textit{both} the query and documents \cite{Dalton2014, Li2015} to expand queries. Apart from entity linking, knowledge bases have also been used for query expansion in several related tasks, such as document filtering \cite{Xie2015} and blog search \cite{Elsas2008, Weerkamp2009}.

%Dalton and Dietz \cite{Dalton2014}, for example, use a number of sources, including entity-annotated queries and pseudo-relevance feedback (PRF) on entity-annotated document collections. Their process is underpinned by entity links produced by their own KB Bridge system \cite{Dalton2013}. Similarly, Xiong and Callan expand queries using pseudo-relevance feedback with the descriptions and categories of Freebase objects linked to a retrieved document in the FACC1\footnote{http://lemurproject.org/clueweb09/FACC1/} annotation data. They employ a supervised machine learning model to select expansion terms, trained on the basis of the terms' influence on the ranking scores of relevant documents \cite{Xiong2015}. Others, including \Li et al. expand queries using the Markov random field model for information retrieval using entity annotations in both the queries and the documents \cite{Li2015}. Other research has investigated query expansion with entities for closely related tasks, like filtering \cite{Xie2015}, entity annotations in the \textit{query}, rather than document \cite{Xu2009, Brandao2014}

@@

\begin{flalign}\label{equation.ql-dirichlet}
\hat{P}(w | \theta_D) = \frac{c(w, D) + \mu \hat{P}(w | C)}{|D| + \mu} 
\end{flalign}

@@

\subsection{Linking with Document Pseudo-Queries}\label{section.linking.queries}
%The KL-divergence retrieval model measures the distance between the query language model, $\theta_{D_Q}$ here, and the document language model, $\theta_E$, as follows:
%
%\begin{flalign*}
%	score(E,D_Q) 	&= -KL(\theta_{D_Q}||\theta_E) \\
%					&\rankeq \sum_{w \in V} P(w|\theta_{D_Q}) \log P(w|\theta_E)
%&\end{flalign*}
%
To find candidate entities to ``link" to a given document $D$, we begin by treating the text of $D$ as a pseudo-query which we pose against a collection of entities $C_E$.  To transform a document into a pseudo-query we apply two transformations.  First we remove all terms from $D$ that appear in the standard Indri stoplist\footnote{http://www.lemurpro ject.org/stopwords/stoplist.dft}.  Next, we prune our pseudo-query by retaining only the $0 < k \leq |D|$ most frequent words in the stemmed text of $D$.  The integer variable $k$ is a parameter that we choose empirically.  Let $Q_D$ be the pseudo-query for $D$, consisting of the text of $D$ after our two transformations.

@@

We now have a ranked list of tuples $\{(E_1, \pi_1)$, $(E_2, \pi_2)$, $...$, $(E_N, \pi_N)\}$ relating knowledge base entry $E_i$ to $D$ with log-probability $\pi_i$. We take the top $n$ entries where $0 \leq n \leq N$. We call these top entries $\mathcal{E}_D$ and designate them as our knowledge base links for $D$.  Finally, we exponentiate each $\pi_i$ and normalize our entity scores so they sum to 1 over the $n$ retained entities.  Assuming a uniform prior over entities, we now have a probability distribution over our $n$ retained entities: $P(E | D)$.

@@

We test our approach using the TREC 2004 robust topics. These 250 topics are used with data from TREC disks 4 and 5. In addition, we use the AP newswire collection from TREC disks 1 and 2 with topics 101-200.

For comparison, we also report the results of our model using entity links produced by Apache Stanbol\footnote{http://stanbol.apache.org/}. We process the test collections with Stanbol's default ``enhancer,'' which supplies entity links and corresponding confidence scores. We select the link with the highest confidence for each mention. These links form the set of document links, and their confidence scores may be normalized to provide an estimate of $P(E|D)$.

@@

We produce three runs per collection:
\begin{itemize}
	\item \textit{baseline-ql}, a baseline query likelihood run
%	\item \textit{baseline-rm3}, a baseline RM3 run
	\item \textit{kb-ql}, incorporating knowledge base links using Eq. \ref{eq.ql-and-entities}
	\item \textit{stanbol-ql}, which uses Stanbol entity annotations in place of our document-level links.
%	\item \textit{kb-rm3}, incorporating knowledge base links using an initial RM3 retrieval
\end{itemize}

For the \textit{kb} and \textit{stanbol} runs, we retrieve the top 1000 documents per query based on the default Indri query likelihood implementation. We then re-rank these documents by incorporating their knowledge base links as described in Section \ref{section.model}.

@@

\begin{table}[htbp]
\centering
\begin{tabular}{|c|l|c|c|} \hline
& {\bf Run} & {\bf $\lambda$} & {\bf MAP} \\ \hline
\multirow{3}{*}{Robust} & {\it baseline-ql} & -- & 0.2183 \\ \cline{2-4}
& {\it kb-ql} & 0.9 & {\bf 0.2275} \\ \cline{2-4}
& {\it stanbol-ql} & 0.1 & 0.1928 \\ \hline \hline 
\multirow{3}{*}{AP} & {\it baseline-ql} & -- & 0.2346 \\ \cline{2-4}
& {\it kb-ql} & 0.9 & {\bf 0.2515} \\ \cline{2-4}
& {\it stanbol-ql} & 0.1 &  \\ \hline
\end{tabular}
\caption{The top-scoring runs and baselines by MAP. Bolded values indicate statistically significant improvements.}
\label{table.performance}
\end{table}

\begin{figure}
\centering
\begin{subfigure}{.5\columnwidth}
\centering
\includegraphics[width=\textwidth]{figures/sweep-ql.pdf}
\end{subfigure}%
\begin{subfigure}{.5\columnwidth}
\centering
\includegraphics[width=\textwidth]{figures/sweep-stanbol.pdf}
\end{subfigure}
\caption{Sweeps over values of $\lambda$ for robust (red) and AP (blue) for (a) \textit{kb} and (b) \textit{stanbol} runs.}
\label{figure.sweeps-ql}
\end{figure}

Retrieval performance of the top-scoring knowledge base runs and baselines are shown in Table \ref{table.performance}. Bolded mean average precision (MAP) scores are greater than the baseline run with statistical significance at $p < 0.001$ using a paired t-test. Note that baselines correspond to $\lambda = 0.0$.

Though Table \ref{table.performance} shows only the top-scoring values of $\lambda$, performance for \textit{kb} runs improved over the baseline for all values of $0.0 < \lambda < 1.0$, as shown in Figure \ref{figure.sweeps-ql}(a). All improvements were statistically significant with at least $p < 0.05$. We may explain the sudden drop in MAP at $\lambda=1.0$ by imagining a scenario in which two documents link to the same set of entities though only one of the documents is a good match for the query. While the entities provide the best data for query likelihood estimation, the loss of this document-specific quantity may be the cause of this drop. Further investigation is needed on this subject.

Notably, our model did not perform well when Stanbol annotations were used as a source of knowledge base links, as seen in Figure \ref{figure.sweeps-ql}(b); in fact, MAP decreased monotonically as more weight was given to the entity model. We attribute this shortcoming to the overfitting of the Stanbol annotations to specific entity mentions in the text, rather than to concepts relevant to the document.

@@

The results indicate that our approach for constructing knowledge base links between documents and Wikipedia produces useful data for document retrieval purposes. Our simple document expansion model that incorporates these links performs well compared to a query likelihood baseline. These outcomes support our argument that a ``bag of links'' to a knowledge base can provide helpful information for a document retrieval task. Further, the poor performance of our model using more traditional mention-to-entity links indicates that, not only are document-to-entity links more efficient to produce, they also connect more useful knowledge base entries for the purposes of document retrieval.
